Fast Bayesian Inference for Exoplanet Discovery in Radial Velocity Data

### 2014-12-15 Paper/ms.tex
Working on the fake data section

\section{Demonstration on Simulated Data}
To test our proposed methodology, we generated a simulated dataset for a
system with $N=5$ planets. The dataset was ``inspired by'' the $\nu$ Oph
dataset (Section~\ref{sec:nu_oph}), and contains two large signals with
periods of 531 and 3180 days, whose semi-amplitudes are
322 ms$^{-1}$ and 214 ms$^{-1}$ respectively. The other three planets have
much lower semi-amplitudes, ranging from 1-9 ms$^{-1}$. The standard deviation
for the noise in the data was 5 ms$^{-1}$, so some of these low-amplitude
signals should be detectable. The simulated data is shown in
Figure~\ref{fig:fake_data}, along with the true radial velocity curve $m(t)$
that was used to generate the data.

\begin{figure}
\includegraphics[scale=0.4]{Figures/fake_data.eps}
\caption{A simulated dataset that ``resembles'' the $\nu$ Oph dataset, which
we used to test our methodology. The dominant signal is from two large
planets with periods periods of 531 and 3180 days and semi-amplitudes of
322 ms$^{-1}$ and 214 ms$^{-1}$ respectively. There are also three
much smaller planets which contribute small additional effects to the data.
\label{fig:fake_data}}
\end{figure}

\begin{figure}
\includegraphics[scale=0.4]{Figures/fake_data_N.eps}
\caption{\label{fig:fake_data_N}}
\end{figure}

\begin{figure}
\includegraphics[scale=0.4]{Figures/fake_data_periods.eps}
\caption{\label{fig:fake_data_periods}}
\end{figure}

@@

The posterior distribution for the logarithms of the periods is shown in
Figure~\ref{fig:nu_oph_periods}. Because of the label-switching degeneracy,
the posterior samples for all periods were combined to make this figure, which
shows three prominent peaks. Defining the log-periods by
$S_i =  \log_{10}\left[T/(\textnormal{1 day})\right]$,
Figure~\ref{fig:nu_oph_periods} is a Monte Carlo representation
of the mixture distribution
\begin{eqnarray}
f(S) &=& \sum_{N=0}^{10} p(N | D)\sum_{i=1}^N p(S_i | N, D).
\end{eqnarray}
The two rightmost peaks are the commonly accepted periods, and the two next
most prominent peaks have periods of 36.12 $\pm$ 0.04 days and
75.57 $\pm$ 0.76 days.
The probability
at least one of these exists (the fraction of our samples which
have one of these signals) is 79\%.
Given that they exist, their amplitudes are low, around 10-15 metres per second,
which we note is above the noise level. We also found that the 75.7 $\pm$ 1.8 day
period prefers an eccentricity of about $0.37 \pm 0.14$.
An example model fit to the data is shown in Figure~\ref{fig:nuoph}.

>>>>

\section{Demonstration on Simulated Data}\label{sec:fake_data}
To test our proposed methodology, we generated a simulated dataset for a
system with $N=5$ planets. The dataset was ``inspired by'' the $\nu$ Oph
dataset (Section~\ref{sec:nu_oph}), and contains two large signals with
periods of 531 and 3180 days, whose semi-amplitudes are
322 m s$^{-1}$ and 214 m s$^{-1}$ respectively. The other three planets have
much lower semi-amplitudes, ranging from 1-9 ms$^{-1}$. The standard deviation
for the noise in the data was 5 m s$^{-1}$, so some of these low-amplitude
signals should be detectable. The simulated data is shown in
Figure~\ref{fig:fake_data}, along with the true radial velocity curve $m(t)$
that was used to generate the data.

\begin{figure}
\includegraphics[scale=0.4]{Figures/fake_data.eps}
\caption{A simulated dataset that ``resembles'' the $\nu$ Oph dataset, which
we used to test our methodology. The dominant signal is from two large
planets with periods periods of 531 and 3180 days and semi-amplitudes of
322 m s$^{-1}$ and 214 m s$^{-1}$ respectively. There are also three
much smaller planets which contribute small additional effects to the data.
\label{fig:fake_data}}
\end{figure}

We ran the MCMC on the simulated dataset to obtain the posterior distribution
for the model parameters. The posterior distribution for $N$, the number of
planets, is shown in Figure~\ref{fig:fake_data_N}. The true number of planets,
5, is also the most probable value according to the posterior distribution,
simply by chance. The posterior distribution also suggests that $N=4$ has
substantial probability.
The posterior distribution for the periods $\{P_i\}_{i=1}^N$ is shown in
Figure~\ref{fig:fake_data_periods}. Because of the label switching degeneracy,
the posterior distribution for each period is identical, so we pooled the
results for all periods. Defining the log-periods by
$S_i =  \log_{10}\left[P_i/(\textnormal{1 day})\right]$,
Figure~\ref{fig:fake_data_periods} is a Monte Carlo representation
of the mixture distribution
\begin{eqnarray}
f(S) &=& \sum_{N=0}^{10} p(N | D)\sum_{i=1}^N p(S_i | N, D).
\end{eqnarray}
The posterior distribution for the periods, shown in
Figure~\ref{fig:fake_data_periods}, shows that four of the true periods were
recovered, with probability close to 1. One period which was actually present
was not detected because it had a very small amplitude. There is also some
evidence for periods which did not exist, however the posterior probabilities
for these peaks are not close to 1.

\begin{figure}
\includegraphics[scale=0.4]{Figures/fake_data_N.eps}
\caption{The posterior distribution for the number of planets $N$ given the
simulated dataset. The true number of planets was 5, which is also the most
probable value (this is a coincidence and will not occur in general).
\label{fig:fake_data_N}}
\end{figure}

\begin{figure}
\includegraphics[scale=0.4]{Figures/fake_data_periods.eps}
\caption{The posterior distribution for the periods\label{fig:fake_data_periods}}
\end{figure}

@@

The posterior distribution for the logarithms of the periods is shown in
Figure~\ref{fig:nu_oph_periods}. As in Section~\ref{sec:fake_data},
the posterior samples for all periods were combined to make this figure, which
shows three prominent peaks. 
The two rightmost peaks are the commonly accepted periods, and the two next
most prominent peaks have periods of 36.12 $\pm$ 0.04 days and
75.57 $\pm$ 0.76 days.
The probability
at least one of these exists (the fraction of our samples which
have one of these signals) is 79\%.
Given that they exist, their amplitudes are low, around 10-15 metres per second,
which we note is above the noise level. We also found that the 75.7 $\pm$ 1.8 day
period prefers an eccentricity of about $0.37 \pm 0.14$.
An example model fit to the data is shown in Figure~\ref{fig:nuoph}.
